Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-26 bbbq_article.tex
Update and add snps_per_gene_name_processed
@bogeert has his own color

The number of variations
was limited to the first 250 variations per gene (see figure
\ref{fig:snps_per_gene_name_ncbi} for the number of SNPs
within the dbSNP database for these gene names),
resulting in \richel{282250} variations.

A variation needs not always be a SNP,
as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
deletions and tandem repeats.
We select only the variations that result in a SNP for
a single amino acid substitution.

@@

@@

% GvdB: A narrative was missing a bit. Why was this done? 
To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
we reanalyzed published peptide elution studies.
Table \ref{tab:elution} shows the percentage of epitopes derived 
from a TMH
found in MHC-I and MHC-II elution 
studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} respectively), 
for the two topology prediction tools TMHMM and 
PureseqTM (\cite{krogh2001predicting} and \cite{wang2019efficient} respectively). 
Regardless of the prediction, 
at least 100 epitopes were predicted to be derived from a TMH. 
% GvdB: Also add a conclusion at the end of the paragraph.
From these findings, we conclude that PureseqTM and TMHMM 
result in prediction of a similar number of TMHs 
and that peptides derived from TMHs have been identified for both MHC-I and -II.

@@

It is known that an individual's haplotype
influences the dynamics of a pathogen (see,
for example, \cite{eccleston2017host} for HIV) \richel{GvdB: Too vague. How?}.
The bioinformatics pipeline developed in this
study may be interesting to make in silico
predictions of a SARS-CoV-2 infection.
The most prominent example is the (MHC-I)
haplotype HLA-A*24:02,
which is predicted to present four times as much epitopes derived 
from SARS-CoV-2 TMHs than expected by chance.
This may help the immune system to detect a SARS-CoV-2 infection earlier,
and change the infection's dynamics.

@@

\richel{
  GvdB:

  Discussion section needs some work, 
  as here interpretation of the results needs to be given. 
  What is the consequence of the findings? Why is it important.

@@

We assume that epitopes 
derived from TMHs are actually presented by MHC-I and MHC-II.
Our analysis (at subsection \ref{subsec:elution_studies}) 
finds that 1.4\% (for MHC-I) and 4.0% (for MHC-II) 
of the epitopes presented 
overlap with a TMH.
It is unsure if this a true finding or a false positive.
One source of noise would stem from imperfect predictions
by the topology prediction software used. 
We have counteracted this by using two different topology prediction
software and confirm these have similar results.
Two biological sources of noise \richel{GvdB: What do you mean by noise?}
are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion \cite{delong2016pathogenic},
two processes that both result in peptide fragments
that did not originate from genome directly.
Note however, that both elution studies have a
methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less 
by mass spectronomy.

@@

\richel{GvdB: This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
% \paragraph{synonymous mutations}
%
In our evolutionary experiment, 
we removed variations that were synonymous mutations (i.e.
resulted in the same amino acid, from a different genetic code) 
from our analysis.
There is evidence, however, that these synonymous mutations
do have an effect and may even be evolutionary selected 
for (\cite{hunt2009silent}).
As the possible effect of synonymous mutations is ignored by our
topology prediction software, we do so as well.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. 
Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
for allowing us to read paywalled articles while working from home.
\richel{GvdB: Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}

@@

>>>>

The number of variations
was limited to the first 250 variations per gene (see figure
\ref{fig:snps_per_gene_name_ncbi} for the number of SNPs
within the dbSNP database for these gene names),
resulting in 60,683 variations at the protein level.
A variation needs not always be a SNP,
as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
deletions and tandem repeats.
We select only the variations that result in a SNP for
a single amino acid substitution, resulting in 38,882 SNPs.

@@

\richel{I suggest to move this section to the Supplementary Materials}

@@

\geert{A narrative was missing a bit. Why was this done?}
To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
we reanalyzed published peptide elution studies.
Table \ref{tab:elution} shows the percentage of epitopes derived 
from a TMH
found in MHC-I and MHC-II elution 
studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} respectively), 
for the two topology prediction tools TMHMM and 
PureseqTM (\cite{krogh2001predicting} and \cite{wang2019efficient} respectively). 
Regardless of the prediction, 
at least 100 epitopes were predicted to be derived from a TMH. 
\geert{Also add a conclusion at the end of the paragraph.}
From these findings, we conclude that PureseqTM and TMHMM 
result in prediction of a similar number of TMHs 
and that peptides derived from TMHs have been identified for both MHC-I and -II.

@@

It is known that an individual's haplotype
influences the dynamics of a pathogen (see,
for example, \cite{eccleston2017host} for HIV) 
\geert{Too vague. How?}.
The bioinformatics pipeline developed in this
study may be interesting to make in silico
predictions of a SARS-CoV-2 infection.
The most prominent example is the (MHC-I)
haplotype HLA-A*24:02,
which is predicted to present four times as much epitopes derived 
from SARS-CoV-2 TMHs than expected by chance.
This may help the immune system to detect a SARS-CoV-2 infection earlier,
and change the infection's dynamics.

@@

\geert{
  Discussion section needs some work, 
  as here interpretation of the results needs to be given. 
  What is the consequence of the findings? Why is it important.

@@

We assume that epitopes 
derived from TMHs are actually presented by MHC-I and MHC-II.
Our analysis (at subsection \ref{subsec:elution_studies}) 
finds that 1.4\% (for MHC-I) and 4.0% (for MHC-II) 
of the epitopes presented 
overlap with a TMH.
It is unsure if this a true finding or a false positive.
One source of noise would stem from imperfect predictions
by the topology prediction software used. 
We have counteracted this by using two different topology prediction
software and confirm these have similar results.
Two biological sources of noise \geert{What do you mean by noise?}
are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion \cite{delong2016pathogenic},
two processes that both result in peptide fragments
that did not originate from genome directly.
Note however, that both elution studies have a
methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less 
by mass spectronomy.

@@

\geert{This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
% \paragraph{synonymous mutations}
%
In our evolutionary experiment, 
we removed variations that were synonymous mutations (i.e.
resulted in the same amino acid, from a different genetic code) 
from our analysis.
There is evidence, however, that these synonymous mutations
do have an effect and may even be evolutionary selected 
for (\cite{hunt2009silent}).
As the possible effect of synonymous mutations is ignored by our
topology prediction software, we do so as well.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. 
Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
for allowing us to read paywalled articles while working from home.
\geert{Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
  \caption{
    Distribution of the number of protein variations and SNPs per gene name processed.
  }
  \label{fig:snps_per_gene_name_processed}
\end{figure}
